---
title: 'Assigment #1'
author: "Ramiro Casó - Incae Business School"
date: "6/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

```{r 9_load_packages, include=FALSE}
rm(list=ls())
library(tidyverse)
library(lme4)
library(broom.mixed)
library(readr)
```

## Importing the data 
```{r}
data_wide <- read_csv("data_wide.csv")
View(data_wide)
```
## Task 1 
Transpose the data from wide form to long form. I also preview the first 10 rows of the new pivoted data set

```{r}
data_long <- data_wide %>%
  pivot_longer(cols = c(y1,y2,y3,y4,y5,x1,x2,x3,x4,x5),
   names_to = c(".value", "Group"),
   names_pattern = "(.)(.)")

head(data_long,10)
```

## Task 2: 
Plot the data in aggregate and describe what you see 

```{r}
ggplot(data = data_long, mapping = aes(x = x, y = y)) +
  geom_point() +
  geom_smooth(method = "lm", na.rm = T, col = "black", se = F) +
  theme(legend.position = "top")
```
What we can see from the data is a slightly negative relationship between x and y. 

## Task 3: 
Plot the data by group to see if there is any evidence of Simpson's paradox. What do you think after looking at this?

In order to plot the data by group, I neet to create a random intercep model first and extract the coefficients. 

```{r}
# First, I need create a fit random intercepts model
intercepts_model <- lmer(y ~ 1+ x + (1 | Group), data = data_long) 

# Then I extract the group's coefficients. 
model_coefs <- coef(intercepts_model)$Group %>% 
  rename(Intercept = `(Intercept)`, Slope = x) %>% 
  rownames_to_column("Group")

# and here we can see coefficients
model_coefs
```

Once I have the coefficients, I need to join them to the original data set. I will create a second data set named `data_long_rani` for that. 

```{r}
data_long_rani <- left_join(data_long, model_coefs, by = "Group")
```

Once I have extracted the coefficients and joined them to my original data set, I can plot it an see if there is evidence of Simpson's Paradox. 

```{r}
model_coef_plot <- ggplot(data = data_long_rani, 
       mapping = aes(x = x, 
                     y = y, 
                     colour = Group)
       ) +
  geom_point(na.rm = T, alpha = 0.5) +
  geom_abline(aes(intercept = Intercept, 
                  slope = Slope,
                  colour = Group
                  ),
              size = 1.5
              ) +
  theme(legend.position = "top")

# see the plot
model_coef_plot
```


## Task 4: 
Fit a pooled model ignoring the hierarchical structure of the data (using the lm function in r)

## Task 5: 
Fit a multilevel model with random intercepts for each group in the data (use the lmer function in r). What changes in the parameters between the pooled and multilevel models?

## Task 6: 
Fit a model with random intercepts and slopes for each group. 

## Task 7: 
Fit a model with crossed random effects (both intercepts and slopes) for both group and id.

## Task 8: 
Because the models in Tasks 5, 6, and 7 are all nested, use a Chi-square test on model deviance to assess whether the more complicated models (i.e., those in Tasks 6 and 7) provide better model fits. To do this, it would be helpful to use the "anova" function in r.

## Task 9: 
[Not graded because there is a chance you won't figure it out in r] get the predicted values (i.e., y_hat) for the model from Task 7. Then plot x and the predicted y's for the first 5 id's in the dataset also accounting for the 5 groups. Here is the code for this last task:

