---
title: 'Assigment #1'
author: "Ramiro Casó - Incae Business School"
date: "6/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

```{r 9_load_packages, include=FALSE}
rm(list=ls())
library(tidyverse)
library(lme4)
library(broom.mixed)
library(readr)
```

## Importing the data 
```{r}
data_wide <- read_csv("data_wide.csv")
View(data_wide)
```
## Task 1 
Transpose the data from wide form to long form. I also preview the first 10 rows of the new pivoted data set

```{r}
data_long <- data_wide %>%
  pivot_longer(cols = c(y1,y2,y3,y4,y5,x1,x2,x3,x4,x5),
   names_to = c(".value", "Group"),
   names_pattern = "(.)(.)")

head(data_long,10)
```

## Task 2: 
Plot the data in aggregate and describe what you see 

## Task 3: 
Plot the data by group to see if there is any evidence of Simpson's paradox. What do you think after looking at this?

## Task 4: 
Fit a pooled model ignoring the hierarchical structure of the data (using the lm function in r)

## Task 5: 
Fit a multilevel model with random intercepts for each group in the data (use the lmer function in r). What changes in the parameters between the pooled and multilevel models?

## Task 6: 
Fit a model with random intercepts and slopes for each group. 

## Task 7: 
Fit a model with crossed random effects (both intercepts and slopes) for both group and id.

## Task 8: 
Because the models in Tasks 5, 6, and 7 are all nested, use a Chi-square test on model deviance to assess whether the more complicated models (i.e., those in Tasks 6 and 7) provide better model fits. To do this, it would be helpful to use the "anova" function in r.

## Task 9: 
[Not graded because there is a chance you won't figure it out in r] get the predicted values (i.e., y_hat) for the model from Task 7. Then plot x and the predicted y's for the first 5 id's in the dataset also accounting for the 5 groups. Here is the code for this last task:

